# Data visualization in ggplot


## Installing and loading the required packages

```{r, eval=FALSE}
install.package("ggplot2")
install.package("plotly")
```


```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(plotly)
```

Disable the scientific notation

```{r}
options(scipen = 999)
```




Import the German dataset

```{r}
german_data_clean <- read_csv("data/german_data_clean.csv")
```


## Normal plots vs ggplot in R

```{r}
#R
plot(german_data_clean$age_years, german_data_clean$credit_amount)
```

```{r}
plot_1 <- ggplot(data = german_data_clean,
                 aes(x = age_years, y = credit_amount)) +
          geom_point()
```

```{r}
plot_1
```

```{r}
plot_1 + geom_point()
```

```{r}
plot_1 + geom_smooth(model = lm)
```



```{r}
plot_1 + geom_rug(sides = "bl")
```


## Different types of visualizations with ggplot

10 bins

```{r}
plot_2 <- ggplot(german_data_clean, aes(credit_amount))
plot_2 + geom_histogram(bins = 10)  
```

Bin width of 1000

```{r}
plot_2 + geom_histogram(binwidth  = 1000)
```


Customizing the colours of your plot

```{r}
plot_2 + geom_histogram(binwidth = 1000, 
                        color="orange", 
                        fill="darkblue")
```


Add fixed vertical lines to show add information about the distribution

```{r}
plot_2 + geom_density(kernel = "gaussian", fill = "gray") +
  geom_vline(aes(xintercept=mean(credit_amount)),
            color="blue", size=1, alpha = 0.4) +
  geom_vline(aes(xintercept=mean(credit_amount) + sd(credit_amount)),
            color="red", size=0.5, alpha = 0.4) +
  geom_vline(aes(xintercept=mean(credit_amount) - sd(credit_amount)),
            color="red", size=0.5, alpha = 0.4) +
  geom_vline(aes(xintercept=median(credit_amount)),
            color="green", size=4, alpha = 0.4) 
```

Scatterplots with multiple dimensions


```{r}
plot_3 <- ggplot(german_data_clean, aes(x = purpose, y = age_years))
plot_3 + geom_boxplot() 
```

```{r}
plot_3 + geom_violin()
```

```{r}
plot_3 + geom_jitter()
```


```{r}
plot_4 <- ggplot(german_data_clean, aes(x = personal_status_sex, y = credit_amount))
plot_4 + geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 250)
```

```{r}
plot_4 <- ggplot(german_data_clean, aes(x = personal_status_sex, y = credit_amount))
plot_4 + geom_boxplot()
```



```{r}
plot_5 <- ggplot(german_data_clean, aes(x = age_years, y = credit_amount)) 
plot_5 + geom_point()
```


## Using multiple dimensions in ggplot


```{r}
plot_5 <- ggplot(german_data_clean, aes(x = age_years, y = credit_amount)) 
plot_5 + geom_point(aes(size = duration_months, colour = purpose, shape = personal_status_sex))
```


Finalizing the visualisation with titles


```{r}
plot_5 <- plot_5 + geom_point(aes(size = duration_months, colour = purpose, shape = personal_status_sex)) +
        ggtitle("Purposes for loans?") +
        xlab("Age") +
        ylab("Amount") +
        theme_minimal()
        
```


## Making visualisations interactive with plotly


```{r}
library(plotly)
ggplotly(plot_5)
```



Another example

```{r}
plot_6 <- ggplot(german_data_clean, aes(purpose, credit_amount)) +    
  geom_bar(aes(fill = personal_status_sex), position = "dodge", stat="identity")

ggplotly(plot_6)
```


Adding more customization


```{r}
plotly_plot_6 <- plot_6 + geom_bar(aes(fill = personal_status_sex), position = "dodge", stat="identity")
ggplotly(plotly_plot_6)
```

```{r}
plotly_plot_5 <- plot_5 + geom_point(aes(size = duration_months, colour = purpose, shape = personal_status_sex)) +
        ggtitle("Purposes for loans") +
        xlab("Age") +
        ylab("Amount") +
        theme(
        panel.grid.major.x = element_line(colour = "#eeeeee"),
        panel.grid.minor.x = element_line(colour = "#000000"),
        panel.background = element_rect(fill = NA)
        )

ggplotly(plotly_plot_5)
```


## Experimenting multiple ways to display your data

```{r}
plot <- ggplot(german_data_clean, aes(x = savings, y = purpose))  + 
               geom_jitter(aes(color = factor(response), shape = personal_status_sex, size = credit_amount), alpha = 0.6) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) #+
  theme_minimal()

ggplotly(plot)

```

Credit history vs. credit amount


```{r}
ggplot() + 
  geom_density(data=german_data_clean, 
               aes(x=credit_amount, group=job, fill=job),alpha=0.5, adjust=2) 
```


### Using different themes

```{r}
library(ggplot2)
theme_set(theme_classic())

# Histogram on a Categorical variable
plot <- ggplot(german_data_clean, aes(purpose))
plot + geom_bar(aes(fill=job), width = 0.5) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Purpose per job")
```



```{r}
german_data_clean <- german_data_clean %>%
  mutate(geslacht = ifelse(personal_status_sex != "female : divorced/separated/married", "man", "vrouw"))

library(ggthemes)
g <- ggplot(german_data_clean, aes(purpose, credit_amount))
g + geom_boxplot(aes(fill=factor(geslacht))) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) 
```

```{r}
german_data_clean <- german_data_clean %>%
  mutate(geslacht = ifelse(personal_status_sex != "female : divorced/separated/married", "man", "vrouw"))

library(ggthemes)
g <- ggplot(german_data_clean, aes(purpose, credit_amount))
g + geom_boxplot(aes(fill=factor(savings))) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) 
```






