# Using the dplyr package in R for data wrangling


## Installing and loading the `dplyr` package


Installing the dplyr package:

```{r, eval = FALSE}
install.packages('dplyr')
```


Using the `library()` to open our package

```{r}
library(dplyr)
```

## Creating subsets of a data frame with dplyr

Base R:

```{r}
older_30 <- data[data$age_years >= 30,]
older_30 <- older_30[, c("age_years", "personal_status_sex", "credit_amount")]
older_30 <- older_30[sort(older_30$age_years),]
```

dplyr:

```{r}
older_30_2 <- data %>%
  filter(age_years >= 30) %>%
  select(age_years, personal_status_sex, credit_amount) %>%
  arrange(age_years)

head(older_30_2)
```

Examples:

* Only **divorsed** men 
* Minimum age of 18
* Loans for: `"car (new)"`, `"radio/television"` or `"furniture/equipment"` :
* Only select the following columns:
  * `"age_years"`
  * `"duration_months"`
  * `"credit_amount"`
  * `"purpose"`
  * `"response"`
* Sorted on `"credit_amount"` from high to low (descending)

Let's see how we can create this subset with using the `dplyr` functionality:

```{r}
selected_purposes <- c("car (new)", "radio/television", "furniture/equipment")

subset_1 <- data %>%
  select(age_years, duration_months, credit_amount, purpose, personal_status_sex, response) %>%
  filter(personal_status_sex == "male   : divorced/separated") %>%
  filter(age_years >= 18) %>%
  filter(purpose %in% selected_purposes) %>%
  arrange(desc(credit_amount))
```


Check out the subset:

```{r}
subset_1
```

## Combining dplyr for creating visualisations for specific subsets


```{r}
hist(subset_1$credit_amount, breaks = 30)
```

```{r}
pie(subset_1$credit_amount, labels = subset_1$purpose)
```

Now, let's show a pie chart for our subset:

```{r}
subset_1_grouped <- subset_1 %>% 
  group_by(purpose) %>%
  summarize(total_amount = sum(credit_amount))
```

```{r}
pie(subset_1_grouped$total_amount, labels = subset_1_grouped$purpose, main = "Loans and purposes.")
```


Challenge:


We will try to create the following subset: 
* Only women (divorced, married, single)
* All ages
* Only loans for `"education"`, `"retraining"` `"business"`, `"other"`:
* Only the following columns:
  * `"age_years"`
  * `"personal_status_sex"`
  * `"credit_amount"`
  * `"purpose"`
  * `"response"`
* Sorted on `"age_years"` descending


* Visualizing the proportion of women that paid back the loan (`"response"`).


```{r}
selected_purposes <- c("education", "retraining", "business", "other")

subset_2 <- data %>%
  filter(personal_status_sex == "female : divorced/separated/married") %>%
  filter(purpose %in% selected_purposes) %>%
  select(age_years, duration_months, credit_amount, purpose, response) %>%
  arrange(desc(age_years))
```


Check out our subset:

```{r}
head(subset_2)
```

```{r}
barplot(table(subset_2$response))
```

## Creating subsets and aggregations in dplyr


Challange: We want to create two bar charts, one for men and one for women, that show the avarage (mean) loan aount grouped by each different purpose:



```{r}
men_woman <- data %>%
  mutate(gender = ifelse(personal_status_sex != "female : divorced/separated/married", "man", "vrouw")) %>%
  select(gender, purpose, credit_amount) %>%
  group_by(gender, purpose) %>%
  summarise(mean_amount = mean(credit_amount)) 
  
men_woman
```


```{r}
men_set <- men_woman %>% filter(gender=="man") 
women_set <- men_woman %>% filter(gender=="vrouw") 

men_amounts <- men_set$mean_amount
names(men_amounts) <- men_set$purpose

women_amounts <- women_set$mean_amount
names(women_amounts) <- women_set$purpose
```


```{r}
barplot(men_amounts, las=2, col="blue")
```

```{r}
barplot(women_amounts, las=2, col="pink")
```
